---
title: "Candy Analysis"
output: html_notebook
---

## Run The Below Chunk To Run Cleaning Script & Assign Clean Data For Analysis

```{r}

# Loads `here` library to support file path/directory management.
library(here)

# Runs cleaning script.
source(here("cleaning_script/cleaning.R"))

# Reads clean data from .csv and assigns to an object for analysis.
candy_analysis <- read_csv(here("data/clean_data/candy_clean.csv"))

```

## Analysis

### Q1.
What is the total number of candy ratings given across the three years. 
Number of candy ratings, not the number of raters. Don’t count missing values

```{r}

candy_analysis %>% 
  filter(response %in% c("JOY", "DESPAIR", "MEH")) %>% 
  summarise(total_ratings = n())

```

### Q2.
What was the average age of people who are going out trick or treating?

```{r}

candy_analysis %>%
  filter(going_out == "Yes") %>%
  distinct(response_id, .keep_all = TRUE) %>% 
  summarise(avg_age = round(mean(age, na.rm = TRUE), 0))

```


### Q3.
What was the average age of people who are not going trick or treating?

```{r}

candy_analysis %>%
  filter(going_out == "No") %>%
  distinct(response_id, .keep_all = TRUE) %>% 
  summarise(avg_age = round(mean(age, na.rm = TRUE), 0))

```

### Q4.
For each of joy, despair and meh, which candy bar received the most of these 
ratings?

```{r}

candy_analysis %>% 
  filter(response == "JOY") %>% 
  count(candy_type, sort = TRUE) %>% 
  head(1)

```

```{r}

candy_analysis %>% 
  filter(response == "MEH") %>% 
  count(candy_type, sort = TRUE) %>% 
  head(1)

```

```{r}

candy_analysis %>% 
  filter(response == "DESPAIR") %>% 
  count(candy_type, sort = TRUE) %>% 
  head(1)

```


### Q5.
How many people rated Starburst as despair?

```{r}

candy_analysis %>% 
  filter(candy_type == "starburst" & response == "DESPAIR") %>%
  summarise(num_starburst_despair = n())
  
```

## In the below questions the below scoring system will be used;

- Joy: +1
- Meh: 0
- Despair: -1

### Q6.
What was the most popular candy bar by this rating system for each gender 
in the dataset ?

```{r}

# Add Score Column

candy_analysis_with_scoring <- candy_analysis %>%
  mutate(score = case_when(
    response == "JOY" ~ 1,
    response == "MEH" ~ 0,
    response == "DESPAIR" ~ -1
  )
)

# Calculate for Males

candy_analysis_with_scoring %>% 
  filter(gender == "Male") %>% 
  group_by(candy_type) %>% 
  summarise(total_score = sum(score, na.rm = TRUE)) %>% 
  arrange(desc(total_score)) %>% 
  head(1)


```

```{r}

# Calculate for Females

candy_analysis_with_scoring %>% 
  filter(gender == "Female") %>% 
  group_by(candy_type) %>% 
  summarise(total_score = sum(score, na.rm = TRUE)) %>% 
  arrange(desc(total_score)) %>% 
  head(1)

```

### Q7.
What was the most popular candy bar in each year?

```{r}

# 2015

candy_analysis_with_scoring %>%
  filter(year == 2015) %>% 
  group_by(candy_type) %>%  
  summarise(total_yearly_score = sum(score, na.rm = TRUE)) %>% 
  arrange(desc(total_yearly_score)) %>% 
  head(1)            

```

```{r}

# 2016

candy_analysis_with_scoring %>%
  filter(year == 2016) %>% 
  group_by(candy_type) %>%  
  summarise(total_yearly_score = sum(score, na.rm = TRUE)) %>% 
  arrange(desc(total_yearly_score)) %>% 
  head(1)  

```

```{r}

# 2017

candy_analysis_with_scoring %>%
  filter(year == 2017) %>% 
  group_by(candy_type) %>%  
  summarise(total_yearly_score = sum(score, na.rm = TRUE)) %>% 
  arrange(desc(total_yearly_score)) %>% 
  head(1)  

```

### Q8.
What was the most popular candy bar by this rating for people in US, Canada, 
UK, and all other countries?

```{r}

# United States of America

candy_analysis_with_scoring %>% 
  filter(country == "United States Of America") %>% 
  group_by(candy_type) %>% 
  summarise(total_score = sum(score, na.rm = TRUE)) %>% 
  arrange(desc(total_score)) %>% 
  head(1)

```


```{r}

# United Kingdom

candy_analysis_with_scoring %>% 
  filter(country == "United Kingdom") %>% 
  group_by(candy_type) %>% 
  summarise(total_score = sum(score, na.rm = TRUE)) %>% 
  arrange(desc(total_score)) %>% 
  head(1)

```


```{r}

# Canada

candy_analysis_with_scoring %>% 
  filter(country == "Canada") %>% 
  group_by(candy_type) %>% 
  summarise(total_score = sum(score, na.rm = TRUE)) %>% 
  arrange(desc(total_score)) %>% 
  head(1)

```

```{r}

# All Others (excluding United States of America, United Kingdom & Canada)

candy_analysis_with_scoring %>% 
  filter(!country %in% c("Canada", "United Kingdom", "United States Of America")) %>% 
  group_by(candy_type) %>% 
  summarise(total_score = sum(score, na.rm = TRUE)) %>% 
  arrange(desc(total_score)) %>% 
  head(1)

```

