---
title: "Candy Analysis"
output: html_notebook
---

## Run The Below Chunk To Run Cleaning Script & Assign Clean Data For Analysis

```{r}

# Loads `here` library to support file path/directory management.
library(here)

# Runs cleaning script.
source(here("cleaning_script/cleaning.R"))

# Reads clean data from .csv and assigns to an object for analysis.
candy_analysis <- read_csv(here("data/clean_data/candy_clean.csv"))

# Add Score Column
candy_analysis_with_scoring <- candy_analysis %>%
  mutate(score = case_when(
    response == "JOY" ~ 1,
    response == "MEH" ~ 0,
    response == "DESPAIR" ~ -1
  )
)

```

## Analysis

### Q1.
What is the total number of candy ratings given across the three years. 
Number of candy ratings, not the number of raters. Don’t count missing values

```{r}

candy_analysis %>% 
  filter(response %in% c("JOY", "DESPAIR", "MEH")) %>% 
  summarise(total_ratings = n())

```

### Q2.
What was the average age of people who are going out trick or treating?

```{r}

candy_analysis %>%
  filter(going_out_trick_or_treating == TRUE) %>%
  distinct(response_id, .keep_all = TRUE) %>% 
  summarise(avg_age = round(mean(age, na.rm = TRUE), 0))

```

### Q3.
What was the average age of people who are not going trick or treating?

```{r}

candy_analysis %>%
  filter(going_out_trick_or_treating == FALSE) %>%
  distinct(response_id, .keep_all = TRUE) %>% 
  summarise(avg_age = round(mean(age, na.rm = TRUE), 0))

```

### Q4.
For each of joy, despair and meh, which candy bar received the most of these 
ratings?

```{r}

candy_analysis %>% 
  filter(response == "JOY") %>% 
  count(candy_type, sort = TRUE) %>% 
  head(1)

```

```{r}

candy_analysis %>% 
  filter(response == "MEH") %>% 
  count(candy_type, sort = TRUE) %>% 
  head(1)

```

```{r}

candy_analysis %>% 
  filter(response == "DESPAIR") %>% 
  count(candy_type, sort = TRUE) %>% 
  head(1)

```

### Q5.
How many people rated Starburst as despair?

```{r}

candy_analysis %>% 
  filter(candy_type == "starburst" & response == "DESPAIR") %>%
  summarise(num_starburst_despair = n())
  
```

## In the below questions the below scoring system will be used;

- Joy: +1
- Meh: 0
- Despair: -1

### Q6.
What was the most popular candy bar by this rating system for each gender 
in the dataset ?

```{r}

# Calculate for Males

candy_analysis_with_scoring %>% 
  filter(gender == "Male") %>% 
  group_by(candy_type) %>% 
  summarise(total_score = sum(score, na.rm = TRUE)) %>% 
  arrange(desc(total_score)) %>% 
  head(1)

```

```{r}

# Calculate for Females

candy_analysis_with_scoring %>% 
  filter(gender == "Female") %>% 
  group_by(candy_type) %>% 
  summarise(total_score = sum(score, na.rm = TRUE)) %>% 
  arrange(desc(total_score)) %>% 
  head(1)

```

```{r}

# Calculate for Other

candy_analysis_with_scoring %>% 
  filter(gender == "Other") %>% 
  group_by(candy_type) %>% 
  summarise(total_score = sum(score, na.rm = TRUE)) %>% 
  arrange(desc(total_score)) %>% 
  head(1)

```
```{r}

# Calculate for I'd rather not say

candy_analysis_with_scoring %>% 
  filter(gender == "I'd rather not say") %>% 
  group_by(candy_type) %>% 
  summarise(total_score = sum(score, na.rm = TRUE)) %>% 
  arrange(desc(total_score)) %>% 
  head(1)

```

```{r}

# Calculate for NA

candy_analysis_with_scoring %>% 
  filter(is.na(gender)) %>% 
  group_by(candy_type) %>% 
  summarise(total_score = sum(score, na.rm = TRUE)) %>% 
  arrange(desc(total_score)) %>% 
  head(1)

```

### Q7.
What was the most popular candy bar in each year?

#### Function Setup / Testing

The below functions will be used to return the most popular candy bar in each 
year. As the most popular candy for all three years was returned as "any full 
sized candy bar", a second function is used to find the most popular candy bar
in these years, excluding "any full sized candy bar".

```{r}

most_popular_candy_by_year <- function(input_data, input_year){
  
  # Check if `input_year` is character or logical value and set error messages.
  
    if (is.character(input_year)){
      stop("Character input for `input_year` - only numeric values accepted")
      }
    
  
    if (is.logical(input_year)){
      stop("Logical input for `input_year` - only numeric values accepted")
      }
  
  # Check to see if there are any instances of the input year in the data.
  
  if (input_year %in% input_data$year){
     
     most_popular_candy <- input_data %>% 
       filter(year == input_year) %>% 
       group_by(candy_type) %>%  
       summarise(total_yearly_score = sum(score, na.rm = TRUE)) %>% 
       arrange(desc(total_yearly_score)) %>% 
       head(1)
     
  # Return a string containing the most popular candy of that year & it's score.
     
    return(str_c("In ", input_year, " the most popular candy was ", most_popular_candy[1], 
                 " with a score of ", most_popular_candy[2]))
  
  # If the are no instances of the year, return the below message.
     
  } else {
    return("Unfortunately we don't have any candy data for this year!")
  }
}

test_that("Character or Logical input returns an error", {
  expect_error(most_popular_candy_by_year$input_year("2011"))
  expect_error(most_popular_candy_by_year$input_year(TRUE))
  expect_error(most_popular_candy_by_year$input_year(FALSE))
})

```

```{r}
most_popular_candy_by_year_minus_any_full_size <- function(input_data, input_year){
  
  # Check if `input_year` is character or logical value and set error messages.
  
    if (is.character(input_year)){
      stop("Character input for `input_year` - only numeric values accepted")
      }
    
  
    if (is.logical(input_year)){
      stop("Logical input for `input_year` - only numeric values accepted")
      }
  
  # Check to see if there are any instances of the input year in the data.
  
  if (input_year %in% input_data$year){
     
     most_popular_candy <- input_data %>% 
       filter(year == input_year & candy_type != "any full sized candy bar") %>% 
       group_by(candy_type) %>%  
       summarise(total_yearly_score = sum(score, na.rm = TRUE)) %>% 
       arrange(desc(total_yearly_score)) %>% 
       head(1)
     
  # Return a string containing the most popular candy of that year & it's score.
     
    return(str_c(most_popular_candy[1], " was the most popular candy in ", input_year, 
                 " exluding any full sized candy bar"))
  # If the are no instances of the year, return the below message.
     
  } else {
    return("Unfortunately we don't have any candy data for this year!")
  }
}

test_that("Character or Logical input returns an error", {
  expect_error(most_popular_candy_by_year_minus_any_full_size$input_year("2011"))
  expect_error(most_popular_candy_by_year_minus_any_full_size$input_year(TRUE))
  expect_error(most_popular_candy_by_year_minus_any_full_size$input_year(FALSE))
})

```

#### Answers

2015:

```{r}

most_popular_candy_by_year(candy_analysis_with_scoring, 2015)
most_popular_candy_by_year_minus_any_full_size(candy_analysis_with_scoring, 2015)

```
2015:

```{r}

most_popular_candy_by_year(candy_analysis_with_scoring, 2016)
most_popular_candy_by_year_minus_any_full_size(candy_analysis_with_scoring, 2016)

```
2017:

```{r}

most_popular_candy_by_year(candy_analysis_with_scoring, 2017)
most_popular_candy_by_year_minus_any_full_size(candy_analysis_with_scoring, 2017)

```

### Q8.
What was the most popular candy bar by this rating for people in US, Canada, 
UK, and all other countries?

```{r}

# United States of America

candy_analysis_with_scoring %>% 
  filter(country == "United States Of America") %>% 
  group_by(candy_type) %>% 
  summarise(total_score = sum(score, na.rm = TRUE)) %>% 
  arrange(desc(total_score)) %>% 
  head(1)

```


```{r}

# United Kingdom

candy_analysis_with_scoring %>% 
  filter(country == "United Kingdom") %>% 
  group_by(candy_type) %>% 
  summarise(total_score = sum(score, na.rm = TRUE)) %>% 
  arrange(desc(total_score)) %>% 
  head(1)

```


```{r}

# Canada

candy_analysis_with_scoring %>% 
  filter(country == "Canada") %>% 
  group_by(candy_type) %>% 
  summarise(total_score = sum(score, na.rm = TRUE)) %>% 
  arrange(desc(total_score)) %>% 
  head(1)

```

```{r}

# All Others (excluding United States of America, United Kingdom & Canada)

candy_analysis_with_scoring %>% 
  filter(!country %in% c("Canada", "United Kingdom", "United States Of America")) %>% 
  group_by(candy_type) %>% 
  summarise(total_score = sum(score, na.rm = TRUE)) %>% 
  arrange(desc(total_score)) %>% 
  head(1)

```

